---
title: 'Peer-graded Assignment: Regression Models Course Project'
author: "Santiago ruiz navas"
date: "11/26/2019"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive summary
This report presents the answers to two research questions:  
    * Is an automatic or manual transmission better for Milles (US) per Gallon (mpg)?  
    * Quantify the mpg difference between automatic and manual transmissions?
These questions will be answered using the dataset mtcars and a three-step process:  
  * Exploratory analysis: review the characteristics of the data and obtain the correlations between the predictor AM and the other predictors  
  * Nested models: Nest various models and evaluate them using ANOVA and Shapiro tests  
  * Inference: Calculate for the 95% interval of the coefficient corresponding to am in the selected model    
 In conclusion:  
  * The use of automatic or manual transmission **does not have a significant effect on mpg**.  
  * The average difference quantity of mpg related to using manual over automatic transmission was **1.301 with a 95% interval of -2.46 and 5.06**.  
  * The present analysis has the limitation that it assumes constant variance of the values of am to mpg and that the data follow normal distributions. The two previous assumptions might not be real in the case of the mtcars dataset. Therefore, the conclusions from the inference analysis might be wrong. 
  
## Exploratory analisys 
 Heads of data
 ```{r expA_summary,echo=T}
library(ggplot2)
data(mtcars)
head(mtcars)
```
In the boxplot below, it is possible to see the values of mpg for automatic or manual transmission:
```{r fig2, fig.height = 2, fig.width = 2, fig.align = "center"}
p <- ggplot(mtcars, aes(x=as.factor(mtcars$am), y=mtcars$mpg)) + 
  geom_boxplot()+ labs(title="Plot of mpg per automatic or manual transmision", x="AM (A = 1, M = 0)",
                       y = "MPG")
p # in mpgs when using automatic vs manual
```
<br/>
From the graph above, it is possible to see a difference between using auotmatic or manual transmission and the value of MPG. However, one question remains, is this difference significant regarding the other variables in the dataset?

## Nested model analysis
The strategy for model selection is nested analysis. I will check the correlations between the predictor AM and the other variables of the dataset, and in the order of the magnitude of their correlations, will create nested models and evaluate them. The justification for this correlation-based-policy is that the more correlated the other predictors are to AM, the more impact they will have on its regressed coefficient. Furthermore, the criteria to select a model is based on how "strong" (p-value) they pass the ANOVA and Shapiro tests.  
  * The evaluation of a model consists of running ANOVA and Shapiro tests.  
  * The "strength" of a model is proportional to the significance in which it passes the tests. e.g., given mdl1 and mdl2, if mdl 1 passes the ANOVA and Shapiro tests with 95% significance and mdl2 with 90%, then mdl1 is stronger than mdl2. 

### Getting the correlations of am
```{r expA_mvar.cor,echo=T, eval=TRUE}
data("mtcars")
mtcars.cor <- cor(mtcars)
am.cor <- sort(abs(mtcars.cor[,"am"]), decreasing = T)
am.cor
```
The nested models are tested in the order showed by the vector am

### Testing the nested models
Three iterations were made to select a model. However, only the first will be shown because of space constrain 

Create the models 
```{r expA_mvar.nmod,echo=T, eval=TRUE}
mdl.project  <- lm(mpg ~ am, mtcars)
mdl.project1 <- lm(mpg ~ am+gear, mtcars)# gear
mdl.project2 <- lm(mpg ~ am+gear+drat, mtcars)# drat
mdl.project3 <- lm(mpg ~ am+gear+drat+wt, mtcars) #wt
mdl.project4 <- lm(mpg ~ am+gear+drat+wt+disp, mtcars) #disp
mdl.project5 <- lm(mpg ~ am+gear+drat+wt+disp+cyl, mtcars) #cyl
mdl.project6 <- lm(mpg ~ am+gear+drat+wt+disp+cyl+hp, mtcars) #hp
mdl.project7 <- lm(mpg ~ am+gear+drat+wt+disp+cyl+hp+qsec, mtcars) #qsec
mdl.project8 <- lm(mpg ~ am+gear+drat+wt+disp+cyl+hp+qsec+carb, mtcars) #carb
mdl.project9 <- lm(mpg ~ am+gear+drat+wt+disp+cyl+hp+qsec+carb+vs, mtcars) #vs
```
Evaluating saphiro test  
```{r expA_mvar.sapt,echo=T, eval=TRUE}
s <- c(shapiro.test(mdl.project$residuals)$p.value,
  shapiro.test(mdl.project1$residuals)$p.value,  shapiro.test(mdl.project2$residuals)$p.value,
  shapiro.test(mdl.project3$residuals)$p.value,  shapiro.test(mdl.project4$residuals)$p.value,
  shapiro.test(mdl.project5$residuals)$p.value,  shapiro.test(mdl.project6$residuals)$p.value,
  shapiro.test(mdl.project7$residuals)$p.value,  shapiro.test(mdl.project8$residuals)$p.value,
  shapiro.test(mdl.project9$residuals)$p.value)
names(s) <- c("m","m1","m2","m3","m4","m5","m6","m7","m8","m9")
s # m4 is bad
```
The p-values of the Shapiro test show that for m4 the null hypothesis does not hold. Therefore, m4 is not adequate.  
<br/>
Evaluating anova test
```{r expA_mvar.anot,echo=T, eval=TRUE}
a <- anova(mdl.project, mdl.project1,  mdl.project2, mdl.project3, mdl.project4,
           mdl.project5, mdl.project6, mdl.project7, mdl.project8, mdl.project9)
a$`Pr(>F)`[5]# m4 is bad
```
Again m4 does not pass the test. Therefore, the variable added in model 4 (disp) is removed from the analysis. 

### Results of nested models
After the three iterations the model **mpg ~ am + gear + drat + wt + cyl** was selected with ANOVA **p-value of 0.00209** and **Saphiro p-value of 0.2862**. Bellow the results of the three iterations are explained  
  * In the first iteration, all variables were nested. After the evaluation, **the variable disp was removed**.  
  * In the second iteration **hp was removed**.  
  * In the third iteration the model **mpg ~ am + gear + drat + wt + cyl** was selected based on its ANOVA and Shapiro tests results  
  
### Inference 
Calculating the inference interval for the estimated coefficient for am:  
The coefficient for am in the selected model was **1.301 with an interval of -2.46 and 5.06 (mpgs)**  
The interval includes cero. Therefore, the difference between using manual over automatic transmission on mpgs is not significant.  

## Conclusions
  * ANSWER TO QUESTION 1: The interval of the estimated coefficient for am in the selected model was tested at a 95% t-test, and it contained cero. Therefore, when taking into account other variables in the dataset, **am does not have a significant effect on mpg**   * ANSWER TO QUESTION 2: the estimated value of for the increase in mpg when using manual over auto was ** 1.301mpg, with an interval of -2.46 and 5.06 (mpgs)**

## Limitations
In this report, I did three iterations of model evaluation; in a future study, other combinations can be explored to find the best possible model, and with it, re-evaluate the answers to these questions.

## Assumptions
The assumptions to compare binary values using linear regressions are   
  * The values for the variables of mpg and am followed a normal distributions  
  * The variances of the set am concerning mpg were equal  

## ANNEX
### Iteration 2
```{r expA_mvar.iter2.mod,echo=T, eval=FALSE}
mdl.project   <- lm(mpg ~ am, mtcars)
mdl.project1  <- lm(mpg ~ am+gear, mtcars)# gear
mdl.project2  <- lm(mpg ~ am+gear+drat, mtcars)# drat
mdl.project3  <- lm(mpg ~ am+gear+drat+wt, mtcars) #wt
mdl.project10 <- lm(mpg ~ am+gear+drat+wt+cyl, mtcars) #cyl
mdl.project11 <- lm(mpg ~ am+gear+drat+wt+cyl+hp, mtcars) #hp
mdl.project12 <- lm(mpg ~ am+gear+drat+wt+cyl+hp+qsec, mtcars) #qsec
mdl.project13 <- lm(mpg ~ am+gear+drat+wt+cyl+hp+qsec+carb, mtcars) #carb
mdl.project14 <- lm(mpg ~ am+gear+drat+wt+cyl+hp+qsec+vs, mtcars) #vs

s <- c(shapiro.test(mdl.project$residuals)$p.value,                      
       shapiro.test(mdl.project1$residuals)$p.value,shapiro.test(mdl.project2$residuals)$p.value,
       shapiro.test(mdl.project3$residuals)$p.value,shapiro.test(mdl.project10$residuals)$p.value,
       shapiro.test(mdl.project11$residuals)$p.value,shapiro.test(mdl.project12$residuals)$p.value,
       shapiro.test(mdl.project13$residuals)$p.value, shapiro.test(mdl.project14$residuals)$p.value)
names(s) <- c("m","m1","m2","m3","m10","m11","m12","m13","m14")

a <- anova(mdl.project, mdl.project1,
           mdl.project2, mdl.project3,
           mdl.project10,mdl.project11,# do not pass the anova hp
           mdl.project12,mdl.project13,
           mdl.project14           )
# Another iteration removing disp and hp
```
### Iteration 3
```{r expA_mvar.iter3.mod,echo=T, eval=FALSE}
mdl.project   <- lm(mpg ~ am, mtcars)
mdl.project1  <- lm(mpg ~ am+gear, mtcars)# gear
mdl.project2  <- lm(mpg ~ am+gear+drat, mtcars)# drat
mdl.project3  <- lm(mpg ~ am+gear+drat+wt, mtcars) #wt
mdl.project10 <- lm(mpg ~ am+gear+drat+wt+cyl, mtcars) #cyl
mdl.project15 <- lm(mpg ~ am+gear+drat+wt+cyl+qsec, mtcars) #qsec
mdl.project16 <- lm(mpg ~ am+gear+drat+wt+cyl+qsec+carb, mtcars) #carb
mdl.project17 <- lm(mpg ~ am+gear+drat+wt+cyl+qsec+vs, mtcars) #vs

s <- c(shapiro.test(mdl.project$residuals)$p.value,                      
       shapiro.test(mdl.project1$residuals)$p.value,shapiro.test(mdl.project2$residuals)$p.value,
       shapiro.test(mdl.project3$residuals)$p.value,shapiro.test(mdl.project10$residuals)$p.value,
       shapiro.test(mdl.project15$residuals)$p.value,shapiro.test(mdl.project16$residuals)$p.value,
       shapiro.test(mdl.project17$residuals)$p.value)
names(s) <- c("m","m1","m2","m3","m10","m15","m16","m17")

a <- anova(mdl.project,mdl.project1, mdl.project2, mdl.project3,
           mdl.project10, mdl.project15,# do not pass the anova hp
           mdl.project16, mdl.project17
)
```
### Confidence interval for the estimated coeficient in the selected model
```{r expA_mvar.confint,echo=T, eval=FALSE}
mdl.project10.coef.statistics <- summary(mdl.project10)$coef
mdl.project10.coef.statistics.B1.int <- 
  mdl.project10.coef.statistics[2,1] +c(-1,1)*qt(.975, df = mdl.project10$df)*mdl.project10.coef.statistics[2,2]
mdl.project10.coef.statistics.B1.int
```