---
title: 'Peer-graded Assignment: Statistical Inference Course Project - Part 1'
author: "Santiago Ruiz Navas"
date: "11/12/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(gridExtra)
set.seed(1)
# Running the simulations
# the would be to 1000 times calculate the average of forthy samples of the exponential distribution.
theoretical.lambda <- 0.2
theoretical.mean <- 1/theoretical.lambda# 5
mns  <- vector("numeric",1000)
for (i in 1 : 1000) mns[i] = mean(rexp(n = 40,rate = theoretical.lambda))

sample.mean     <- mean(mns)
sample.lambda   <- 1/sample.mean
############################################################################
## first point comparing the theoretical mean against the simulated means.## 
############################################################################
############################################################################
##                   second point comparing variances.                    ## 
############################################################################
sample.var      <- 1/sample.lambda^2
theoretical.var <- 1/theoretical.lambda^2
### analysis, almost the same

exponential <- rexp(n = 1000,rate = theoretical.lambda)
g.exp = ggplot(data.frame(Exponential = exponential), aes(x= Exponential))
g.exp = g.exp+geom_histogram(color = "black", fill = "lightblue", binwidth = 0.05 )

g.exp.var =g.exp +geom_vline(xintercept = theoretical.mean + sqrt(theoretical.var), size = 1, colour = "red")
g.exp.var =g.exp.var +geom_vline(xintercept = theoretical.mean - sqrt(theoretical.var), size = 1, colour = "red")
g.exp.var =g.exp.var +geom_vline(xintercept = theoretical.mean + sqrt(sample.var), size = 1, colour = "blue")
g.exp.var =g.exp.var +geom_vline(xintercept = theoretical.mean - sqrt(sample.var), size = 1, colour = "blue")
g.exp.var =g.exp.var+labs(title = paste("Theoretical(",theoretical.var,")","vs Simulated Variance(",round(sample.var,3),")"),subtitle ="lambda  =  0.2")

g.exp.mean = g.exp+ geom_vline(xintercept = theoretical.mean, size = 1, colour = "red")
g.exp.mean = g.exp+ geom_vline(xintercept = sample.mean, size = 1, colour = "blue")
g.exp.mean =g.exp.mean+labs(title = paste("Theoretical(",theoretical.mean,")","vs Simulated Mean(",round(sample.mean,3),")"),subtitle ="lambda  =  0.2")

g.mean = ggplot(data.frame(means = mns), aes(x= means))
g.mean = g.mean+geom_histogram(color = "black", fill = "lightblue", binwidth = 0.05 )
g.mean = g.mean+geom_vline(xintercept = sample.mean, size = 1, colour = "red")
g.mean = g.mean+labs(title = "Simulated mean",
                     subtitle =paste("u =",sample.mean,"\nsigma =",sample.var))

g.var = ggplot(data.frame(means = mns), aes(x= means))
g.var = g.var+geom_histogram(color = "black", fill = "lightblue", binwidth = 0.05 )
g.var = g.var+geom_vline(xintercept = sample.mean + sqrt(sample.var), size = 1, colour = "red")
g.var = g.var+geom_vline(xintercept = sample.mean - sqrt(sample.var), size = 1, colour = "red")
g.var = g.var+labs(title = "Simulated variance",
                     subtitle =paste("u =",sample.mean,"\nsigma =",sample.var))
#g
### analysis
# The simulated mean converged really close to the theoretical, Therefore, the simulated 40 averages of the exponential distributions drawn in this plot comply with the premises of the CTL. 

############################################################################
##                   third point comparing distributions.                 ## 
############################################################################
exponential <- rexp(n = 1000,rate = theoretical.lambda)
g1 = ggplot(data.frame(Exponential = exponential), aes(x= Exponential))
g1 = g1+geom_histogram(color = "black", fill = "lightblue", binwidth = 0.05 )
g1 = g1+labs(title = "PART 1 Simulation - Graph 1",
             subtitle ="Exponential distribution")

g2 = ggplot(data.frame(Means = mns), aes(x= Means))
g2 = g2+geom_histogram(color = "black", fill = "lightblue", binwidth = 0.05 )
g2 = g2+labs(title = "PART 1 Simulation - Graph 2",
                subtitle ="Distribution of the averages of 40 exponentials")

#grid.arrange(g1, g2, nrow = 1)
### Analysis
# The two graphs show a explicit differences between the distributions. 
# Graph one shows 1000 samples of an exponential distribution with lambda 0.2
# Graph two shows 1000 averages of 40 points of a exponential distribution with lambda 0.2. 
# From these two graphs it is possible to see that graph two resembles an normal distrubution as indicated by the CTL.

#--------------------------------------------------------------------------------------------------------------------
# PART 2 - Basic Inferential Data Analysis Instructions
#--------------------------------------------------------------------------------------------------------------------
data(ToothGrowth)
#summary(ToothGrowth)
#head(ToothGrowth)
#ToothGrowth$dose
# paired measures, differences between two treatments was positive or not
## by supp
data.VC <- subset(ToothGrowth,supp == "VC")
data.OJ <- subset(ToothGrowth,supp == "OJ")

difference <- data.VC$len-data.OJ$len
#t.test(difference)$conf.int

#Given that the samples were taken from the same patients over a 

#t.test(g2, g1, paired = FALSE, var.equal = TRUE)$conf.int
#t.test(data.VC$len, data.OJ$len, paired = FALSE, var.equal = FALSE)
# For paired the two sided t-test at 95% provides an interval -7.57 and 0.17, the interval includes cero. Therefore, when comparing OJ and VC we fail to reject the null hipothesis and we can not conclude wether OJ or VC have an effect on tooth lenght at 95% of confidence inteval.

## by dose
data.dose1 <- subset(ToothGrowth,dose == 0.5)
data.dose2 <- subset(ToothGrowth,dose == 1.0)
data.dose3 <- subset(ToothGrowth,dose == 2.0)

df12 <- data.dose1$len-data.dose2$len
df13 <- data.dose1$len-data.dose3$len
df23 <- data.dose2$len-data.dose3$len

data12 <- t.test(df12, paired = FALSE, var.equal = FALSE)
# all negative, therefore dose 2 increses with respecto to dose 1
# t -6.9
# p-value super small

data13<- t.test(df13, paired = FALSE, var.equal = FALSE)
# all super negative, therefore dose 3 increses with respecto to dose 1
# t = -11.291
# p-value = close to cero
data23 <- t.test(df23, paired = FALSE, var.equal = FALSE)
# t = -4.6
# p-value = 0.00019
# all negative, therefore dose 3 increses with respecto to dose 2
# The dose quantity is proportional to the increase in tooth lenght

# Combining the two analyses it is proposed that whit the current data, we can not say that the type of supp has an effect on tooth grwoth However, it is possible to say that the more doses there are of a treathment the longer the tooth grow. 

```
# OVERVIEW
This document consits in two sections, part 1 corresponding to the simualtion part of the  Statistical Inference Course Project.

# 1. Simulations
## 1.1 calculating the theoretical and simulated means
The theoretical mean was calculated by 1/lambda with lambda = 0.2. 
```{r,theoretical.mean, echo=TRUE}
theoretical.lambda <- 0.2
theoretical.mean <- 1/theoretical.lambda
theoretical.mean
```
The simulated mean was calculated by calculating the mean of 1000 averages of 40 randomly generated samples of the exponential distribution with lambda = 0.2
```{r,theoretical.lambda, echo=TRUE}
set.seed(1)
mns  <- vector("numeric",1000)
for (i in 1 : 1000) mns[i] = mean(rexp(n = 40,rate = theoretical.lambda))
sample.mean     <- mean(mns)
sample.mean
```
## 1.2 Calculating the theoretical and simulated variance
The theoretical variance of an exponential distribution is calcualte with 1/lambda^2
```{r, echo=TRUE}
theoretical.var <- 1/theoretical.lambda^2
theoretical.var
```
The simulated variance is calculated using the simualted mean obtained in point 1 and 1/lambda^2
```{r,mns, echo=TRUE}
sample.mean     <- mean(mns)
sample.lambda   <- 1/sample.mean
sample.var      <- 1/sample.lambda^2
sample.var
```
### 1.2.1 Simulated vs theoretical mean
I present in Figure bellow the distribution of the simulated data. Also two lines, in "blue" a vertical line showing the value of the theoretical mean (5) and in "red" the simulated mean

```{r g.exp.mean, g.mean, echo=FALSE}
g.exp.mean
```

### 1.2.2 Simulated vs theoretical variance
Bellow it is possible to see the values of the theoretical and simulated square root of the variances
```{r g.exp.var, echo=FALSE}
g.exp.var
```

### 1.2.3 Distribution
In the figure bellow the histogram of the 1000 averages of 40 randomly generated samples of the exponential distribution with lambda = 0.2. The distribution shows a bell shaped curved resembling a normal distribution. 
```{r g.mean, echo=FALSE}
g.mean
```

### 1.2.4 Interpretation
The differences between the theoretical and simulated means and variances was close to cero.
Therefore, the simulated 40 averages of the exponential distributions drawn in this plot comply with the premises of the CTL.